Painting with Triangles

### 2014-04-22 paper.tex
Updated boxes
Fix tilde.
move figure.
Increase figure 12 size.
foo

\documentclass[review]{acmsiggraph}

@@

@@

@@

\teaser{
  \setlength{\w}{0.33\textwidth}
  \centering
  \includegraphics[width=\w]{images/rainbow3}
  \includegraphics[width=\w]{images/rainbow3_tri}
  \includegraphics[width=\w]{images/rainbow3_zoom3}
  \caption{\emph{Left:} A vector painting made in the system exhibits both smooth gradients painted by soft brushes as well as infinitely sharp boundaries painted by hard brushes. \emph{Middle:} The painting is represented by a planar triangle mesh. \emph{Right:} Closeup on the bottom of the letter B highlights both the sharp boundaries on the white region and smooth gradient from blue to green in the background.}
  \label{fig:teaser}
}

@@

\begin{abstract}
Although vector graphics offer a number of benefits, conventional vector painting programs
do not allow an artist to use traditional (digital) painting techniques. 
% \note{really?}
We propose a new algorithm that translates a
user's mouse motion into a triangle mesh representation. This triangle mesh can then be composited onto
a canvas containing an existing mesh representation of the previous strokes. 
This representation allows the algorithm to render solid colors and linear
gradients. It also enables painting at any resolution. This gives artists the opportunity to create complex,
multi-scale drawings with gradients while avoiding artifacts.

@@

To transform a user's mouse inputs to a triangle mesh,
the system uses a combination of rasterization and marching squares to
find stroke contours. From the contour it performs a Delaunay triangulation
and then merges this triangulation
of the new stroke with the triangle mesh of the existing canvas. This process yields
a new canvas with the combined strokes.

@@

\begin{figure}[t]
    \centering
        \includegraphics[width=0.5\textwidth]{images/gradientarcs}
    \caption{A close look at a gradient stroke. The blue edges are boundaries where
    the points have gray color arcs as shown. The purple edges are boundaries where
    the points have transparent gray color arcs. This gives triangles connecting the two boundaries
    a linear gradient from gray to white.}
    \label{fig:softarcs}
\end{figure}

@@

\subsection{Converting Strokes to Contours}
A set of contours define the outline of a stroke. For simply connected strokes, those without holes, 
one contour describes the whole stroke. However, strokes with holes, such as a circle or 
figure-eight, will have more than one contour.
Converting a user's mouse motion into a set of contours requires two steps.
First, the user's mouse movements are captured by a set of polygons which can be
rendered to yield a rasterized version of the stroke.
Second, to obtain the contours the stroke is rasterized in black and marching squares~\cite{lorensen1987}
is used to extract iso-contours of 50\% gray. This returns a very dense set of points,
which we reduce through a pruning process to
get a sparse but accurate representation of the contour.
The pruning procedure walks around the contour to form an approximating polygon $P$,
greedily eliding vertices whenever they lie within a small threshold distance of $P$.

@@

\subsection{Converting Contours to Triangles}
The contour contains two important pieces of information. First, it describes all of the points of the stroke.
Second, segments connecting adjacent points describe the boundaries of the stroke. 
A triangle mesh must contain all the points and stay
within the boundaries. 
For example, a concave contour like the outline of the letter C should not contain 
triangles inside the concavity, since such triangles would be outside of the boundary.
These boundaries will prove to be crucial when compositing a stroke onto the canvas.
%
We perform constrained Delaunay triangulation on the given set of points and boundaries using
Shewchuk's implementation~\shortcite{shewchuk2002}, and the output is a mesh that
approximately represents the drawn stroke.

@@

Again, using marching squares these contours are extracted and then pruned. The boundaries and points
derived from the contours are given to Triangle and a mesh is returned. This mesh represents the soft
stroke where no triangle crosses the boundary between the soft and hard regions
of the stroke. Points on the contour of the inner stroke will have color arcs with the stroke color,
while the points on the contour of the outer stroke will have transparent color arcs. This gives
all triangles connecting the inner and outer strokes a linear gradient (again see Figure~\ref{fig:softarcs}).
Figure~\ref{fig:rastertotriangles} shows the process of triangulating both hard and soft strokes.

@@

@@

In the simplest approach, the points and boundaries from the new stroke are added
to the list of points and boundaries already stored on the canvas. All of this data is given to Triangle
to produce a new mesh (see figure~\ref{fig:firststrokes}). The new mesh contains all of the points of the old and new strokes. Since
all of the boundaries are stored and given to Triangle, the stroke edges are preserved. This means that
no triangle can cross the boundary of a stroke. Without this constraint, it would be impossible to correctly color the
triangles since they would straddle color arcs. Once a new mesh is formed the color arcs of the 
new and old points must be computed.

\begin{figure}
    \centering
        \includegraphics[width=.23\textwidth]{images/emptycanvas}
        \includegraphics[width=.23\textwidth]{images/canvaswithonestroke}
    \caption{The figure on the left shows an empty canvas with a number of points and boundaries defined along the edges
             to provide the initial triangulation. The figure on the right shows how a stroke is composited onto
             the empty canvas. Blue edges are boundary edges. Red edges form the triangle mesh with the blue edges.}
    \label{fig:firststrokes}
\end{figure}

\subsection{Determining Colors}
All color compositing is done with regular RGB alpha blending.  The color arcs for new points depend on two values, the color of the stroke and the color
of the canvas at that point. The points on a hard stroke will have two color arcs. One
arc describes the inside of the stroke. The color for this arc is the color of the stroke
composited over the color of the canvas at that point. The other arc describes the outside
of the stroke. The color for this arc is just the color of the canvas at that point. The
points on a soft stroke will each have one color arc. The color arc for points on the inner
contour is the color of the stroke composited over the color of the canvas at that point.
The color arc for points on the outer contour is the color of canvas at that point (because the outer edge of the soft stroke is completely transparent).
This gives any triangle connecting inside and outside contour points a linear gradient
from the stroke color composited over the canvas color to the canvas color.

The color arcs for old points remain the same unless part of the new stroke covers it.
In this case the colors in each arc must be replaced with the stroke color composited over
the old color of the arc. If the new stroke is soft then the color at the location of
the old point must be determined through bilinear interpolation before compositing.

\begin{figure}
    \centering
        \includegraphics[width=.2\textwidth]{images/intersectioncolorarcs1final}
        \includegraphics[width=.2\textwidth]{images/intersectioncolorarcs2final}

@@

When Triangle is asked to triangulate a set of points with intersecting boundary edges, it must
insert a point at the intersection, since without it at least one triangle would have to
cross a boundary. These points are introduced with no color information, but the final colors for 
all of the other points in the mesh have already been determined.

An intersection point lies on two boundaries and therefore has four boundary edges emanating from it. 
To determine the colors of an intersection points requires knowing the colors associated with
other points on the boundaries. For both boundaries, travel both directions to the first point with color data. Note this may not
be the first point on each side since this intersection point may be adjacent to other intersection
points also without color data. For both boundaries, the colors in the color arcs are linearly interpolated. 
This yields four average colors, two from the new stroke boundary and two from the old stroke boundary.
The colors from the new stroke are composited over the colors from the old stroke to yield four new
colors, one for each region defined by the boundary intersections (see figure~\ref{fig:intersectioncolors})

An example of how color arcs are determined for both intersection and regular points see figure~\ref{fig:colorarcdeterminer}.

\subsection{Finding Modified Points}
The above procedure has two important drawbacks. First, it processes many points that do not need
consideration which takes a considerable amount of processing time. Second, since it acts on the whole
canvas, local changes can have global effects. To avoid this problem the algorithm determines
which triangles can remain
and which triangles need to be included in the re-triangulation (see Figure~\ref{fig:modified}). To accomplish this a static grid
is generated and all of the cells which contain a triangle from the new stroke are marked. 
The set of triangles from the old triangulation that intersect these grid squares are found.
The set of edges from these triangles that do not intersect these grid cells form a ring
around the area of our new triangles. The edges in this ring become constraints
in the triangulation. All of the points and constraint edges from the old triangulation are
included as well as the points and constraint edges from the new triangulation. Triangulating
these points and edges gives the new geometry for the modified area and has no effect on any
other part of the canvas. Therefore, the old triangles from the rest of the canvas can be reused. 
The new triangulation combined with the old triangles give a new triangle mesh representation of the entire canvas.

\section{Discussion and Results}
Paintings made with the system can be seen in Figures~\ref{fig:face}, \ref{fig:teaser}, \ref{fig:stop} and \ref{fig:boxes}.  Here we first consider how the complexity of the algorithm grows both in time and geometry.
The results from this analysis and our experience with the system have lead to a number of observations.

@@

Figure~\ref{fig:numtriangles} shows the zooming case grows linearly while the non-zooming case
appears to grow quadratically. This is expected since every new stroke has the chance to overlap
all of the old strokes. Therefore each new stroke does not just add a constant amount of geometry, but 
due to intersections with previous strokes extra intersection points are added. This results in a linear
increase in the amount of geometry per stroke yielding a quadratic growth rate. However, in the zooming case each stroke rarely overlaps and therefore
intersection points are rarely added. In this case the geometry increases by a constant amount per stroke which yields
a linear growth rate.

@@

\subsection{Stability}
While the Triangle library is impressively robust, there are still certain pathological geometries that can cause it to fail.  Specifically, long skinny triangles and vertices that are too close together seem to trigger triangulation failures.  Detecting these situations and fixing them, or even better modifying the geometry construction to avoid them, would solve this problem.  For example, identifying if a new vertex is too close to an existing one, and if so, merging them.  Another example is to subdivide long edges to reduce the incidence of skinny triangles.  The impact these solutions would have on processing time is something that needs to be explored.

\begin{figure}
    \centering
        \includegraphics[width=\columnwidth]{images/grid}
    \caption{An example of how the system determines which triangles to save and which to retriangulate. Blue
    lines represent the boundaries of old strokes, orange lines are the boundaries of the new stroke, red lines are
    the triangles that will be reused, and purple lines define the area that needs to be retriangulated. The purple
    lines are edges of triangles that intersect the new stroke. Since they interesect the new stroke they are no
    longer valid and must be retriangulated. All other triangles on the canvas can be reused.}
    \label{fig:modified}
\end{figure}

@@

\begin{figure}
  \centering
  \begin{subfigure}[b]{0.5\columnwidth}
    \centering
    \includegraphics[width=\columnwidth]{graphs/numtriangles}
    \caption{Number of triangles}
    \label{fig:numtriangles}
  \end{subfigure}%
  \begin{subfigure}[b]{0.5\columnwidth}
    \centering
    \includegraphics[width=\columnwidth]{graphs/cumulativetime}
    \caption{Rendering time}
    \label{fig:timing}
  \end{subfigure}%
  \caption{For paintings of 50 random strokes, the left graph shows the number of triangles in the canvas as of stroke $x$, and the right graph shows the cumulative render time of the painting. The blue points are the non-zooming case, while the red points are the zooming case.}
\end{figure}

@@

% - Switch 6.3 and 6.4
\begin{figure*}
    \centering
        \includegraphics[width=0.49\textwidth]{images/facezoom}
        \includegraphics[width=0.49\textwidth]{images/facezoomtriangles}
    \caption{A drawing by an artist that took ~30 minutes. 9030 triangles.} % \Steve{Can we add number of strokes?}}
    \label{fig:face}
\end{figure*}

\begin{figure*}
    \centering
        \includegraphics[width=0.3\textwidth]{images/stop12}
        \includegraphics[width=0.3\textwidth]{images/stop22}
        \includegraphics[width=0.3\textwidth]{images/stop32}
    \caption{Another painting by an artist, using soft and hard strokes at a variety of scales.  Middle and right images are zoomed in.}
    \label{fig:stop}
\end{figure*}

\begin{figure*}
    \centering
        \includegraphics[width=0.195\textwidth]{images/zoom61}
        \includegraphics[width=0.195\textwidth]{images/zoom62}
        \includegraphics[width=0.195\textwidth]{images/zoom63}
        \includegraphics[width=0.195\textwidth]{images/zoom64}
        \includegraphics[width=0.195\textwidth]{images/zoom65}
        \includegraphics[width=0.195\textwidth]{images/zoom66}
        \includegraphics[width=0.195\textwidth]{images/zoom67}
        \includegraphics[width=0.195\textwidth]{images/zoom68}
        \includegraphics[width=0.195\textwidth]{images/zoom69}
        \includegraphics[width=0.195\textwidth]{images/zoom72}
    \caption{A painting that shows extreme zoom levels.  The final image is at zoom level 524,000.}
    \label{fig:boxes}
\end{figure*}

\section{Conclusion}
The system enables a user to paint like they would in a raster graphics program 
to create vector graphics. The core of the system is a triangle mesh representation of the image.
We have created an algorithm to convert a user's mouse motions into a triangle mesh, and then composite
that mesh on the preexisting mesh containing the previous strokes.

The current implementation provides a baseline painting experience. Future work should focus on
implementing a simplification algorithm to allow the complexity of the drawings to grow. Other
improvements might focus on creating interesting manipulation tools, or other operations that utilize
the underlying triangle mesh structure to make the artists' job easier.

\section*{Acknowledgments}

Omitted for review.

>>>>

% uncomment one the following two lines, first one showing pictures
\documentclass[review]{acmsiggraph}
%\documentclass[review,draft]{acmsiggraph}

@@

% comment out this line to get figures; uncomment for fast latex
%\usepackage[draft]{graphicx}
\usepackage{graphicx}

@@

\setlength{\fboxsep}{0pt}
\setlength{\fboxrule}{1pt}

@@

\teaser{
  \setlength{\w}{0.33\textwidth}
  \centering
  \includegraphics[width=\w]{images/rainbow3}
  \includegraphics[width=\w]{images/rainbow3_tri}
  \includegraphics[width=\w]{images/rainbow3_zoom3}
  \caption{\emph{Left:} A vector painting made in our system exhibits both smooth gradients painted by soft brushes as well as infinitely sharp boundaries painted by hard brushes. \emph{Middle:} The painting is represented by a planar triangle mesh. \emph{Right:} Closeup showing the bottom of the letter B reveals both the sharp boundaries on the white region and a smooth gradient from blue to green in the background.}
  \label{fig:teaser}
}

@@

\begin{abstract}
Although vector graphics offer a number of benefits, conventional vector painting programs
offer only limited support for the traditional painting metaphor.
We propose a new algorithm that translates a
user's mouse motion into a triangle mesh representation. This triangle mesh can then be composited onto a canvas containing an existing mesh representation of earlier strokes. 
This representation allows the algorithm to render solid colors and linear
gradients. It also enables painting at any resolution. This paradigm allows artists to create complex, multi-scale drawings with gradients and sharp features while avoiding pixel sampling artifacts.

@@

To transform a user's mouse inputs to a triangle mesh,
our system uses a combination of rasterization and marching squares to
find stroke contours. From the contour it performs a Delaunay triangulation
and then merges this triangulation
of the new stroke with the triangle mesh of the existing canvas. This process yields
a new canvas with the combined strokes.

@@

\begin{figure}[t]
    \centering
        \fbox{\includegraphics[width=0.45\textwidth]{images/gradientarcs}}
    \caption{A close look at a gradient stroke. The blue edges are boundaries where
    the points have gray color arcs as shown. The purple edges are boundaries where
    the points have transparent gray color arcs. This gives triangles connecting the two boundaries
    a linear gradient from gray to white.}
    \label{fig:softarcs}
\end{figure}

@@

\subsection{Converting Strokes to Contours}

A set of contours define the outline of a stroke. For simply connected strokes, those without holes, 
one contour describes the whole stroke. However, strokes with holes, such as a circle or 
figure-eight, will have more than one contour.
Converting a user's mouse motion into a set of contours requires two steps.
First, the user's mouse movements are captured by a set of polygons which can be
rendered to yield a rasterized version of the stroke.
Second, to obtain the contours the stroke is rasterized in black and marching squares~\cite{lorensen1987}
is used to extract iso-contours of 50\% gray. This returns a very dense set of points,
which we reduce through a pruning process to
get a sparse but accurate representation of the contour.
The pruning procedure walks around the contour to form an approximating polygon $P$,
greedily eliding vertices whenever they lie within a small threshold distance of $P$.

@@

\subsection{Converting Contours to Triangles}
\label{sec:contours}

Now that we have the contour we need to create a triangulation for the stroke. 
First we consider the case of a hard brush, which contains a single contour (\fig{fig:rastertotriangles}-left).
Points on the contour describe the shape of the stroke, while the adjoining segments
describe its boundary.
A triangle mesh must contain all the points and stay
within the boundaries. 
For example, a concave contour like the outline of the letter C should not contain 
triangles inside the concavity, since such triangles would be outside of the boundary.
These boundaries will prove to be crucial when compositing a stroke onto the canvas.
%
We perform constrained Delaunay triangulation on the given set of points and boundaries using
Shewchuk's implementation~\shortcite{shewchuk2002}, and the output is a mesh that
approximately represents the drawn stroke.

@@

Again, using marching squares these contours are extracted and then pruned. The boundaries and points
derived from the contours are triangulated as in \sect{sec:contours}.
The resulting mesh represents the soft
stroke where no triangle crosses the boundary between the soft and hard regions
of the stroke. Points on the contour of the inner stroke will have color arcs with the stroke color,
while the points on the contour of the outer stroke will have transparent color arcs. This gives
all triangles connecting the inner and outer strokes a linear gradient as seen in Figure~\ref{fig:softarcs}.
Figure~\ref{fig:rastertotriangles}-right shows the process for triangulating soft strokes.

@@

\begin{figure}
    \centering
        \fbox{\includegraphics[width=.22\textwidth]{images/emptycanvas}}
        \fbox{\includegraphics[width=.22\textwidth]{images/canvaswithonestroke}}
    \caption{The figure on the left shows an empty canvas with a number of points and boundaries defined along the edges
             to provide the initial triangulation. The figure on the right shows how a stroke is composited onto
             the empty canvas. Blue edges are boundary edges. Red edges form the triangle mesh with the blue edges.}
    \label{fig:firststrokes}
\end{figure}

@@

In the simplest approach, the points and boundaries from the new stroke are added
to the list of points and boundaries already stored on the canvas. 
%
These points and boundaries are all triangulated using constrained Delaunay triangulation
as in \sect{sec:contours}, 
to produce a new mesh (Figure~\ref{fig:firststrokes}).
%
The new mesh contains all of the points of the old and new strokes. 
Since boundary edges form constraints in the triangulation, 
stroke edges are preserved. This means that
no triangle will cross the boundary of a stroke. 
Without this constraint, it would be impossible to correctly color the
triangles since they would straddle color arcs.

\subsection{Determining Colors}

Once a new mesh is formed the color arcs of the 
new and old points must be computed.
%
All color compositing is done with regular RGB alpha blending.  The color arcs for new points depend on two values, the color of the stroke and the color
of the canvas at that point. The points on a hard stroke will have two color arcs. One
arc describes the inside of the stroke, which takes the color of the stroke
composited over the color of the canvas at that point. The other arc describes the outside
of the stroke, which is just the color of the canvas at that point. The
points on a soft stroke will each have one color arc. The color arc for points on the inner
contour is the color of the stroke composited over the color of the canvas at that point.
The color arc for points on the outer contour is the color of canvas at that point (because the outer edge of the soft stroke is completely transparent).
This arrangement produces a linear gradient across the triangles separating the inner
and outer contours.

The color arcs for old points remain unchanged, unless part of the new stroke covers it.
In this case the colors in each arc must be replaced with the stroke color composited over
the old color of the arc. If the new stroke is soft then the color at the location of
the old point must be determined through bilinear interpolation before compositing.

\begin{figure}
    \centering
        \includegraphics[width=.2\textwidth]{images/intersectioncolorarcs1final}
        \hspace{.2in}
        \includegraphics[width=.2\textwidth]{images/intersectioncolorarcs2final}

@@

When the constrained Delaunay triangulation is computed
for a set of points with intersecting boundary edges, it must
insert a new point at the intersection of the edges
(otherwise at least one triangle would have to
cross a boundary). These points are introduced with no color information, 
whereas the final colors for 
all of the other points in the mesh have already been determined.

An intersection point lies on two boundaries and therefore has four boundary edges emanating from it. 
To determine the colors of an intersection points requires knowing the colors associated with
other points on the boundaries. For both boundaries, 
we search both directions to the first point with color data. Note this may not
be the first point on each side, since this intersection point may be adjacent to 
\emph{other} intersection
points also without color data. For both boundaries, the colors in the color arcs are linearly interpolated. 
This yields four average colors, two from the new stroke boundary and two from the old stroke boundary.
The colors from the new stroke are composited over the colors from the old stroke to yield four new
colors, one for each region defined by the boundary intersections (see Figure~\ref{fig:intersectioncolors}).
% XXX
For an example of how color arcs are determined for both intersection and regular points see Figure~\ref{fig:colorarcdeterminer}.

\subsection{Finding Modified Points}

As described so far, our retriangulation procedure has two substantial drawbacks. 
%
First, it processes many points that do not need consideration which is inefficient. 
Second, since it acts on the whole
canvas, local modifications may have global effects. 
%
Instead our approach is to determine
which triangles can remain unchanged
and which triangles need to be included in the re-triangulation (see Figure~\ref{fig:modified}). To accomplish this a static grid
is generated and all of the cells that contain a triangle from the new stroke are marked. 
The set of triangles from the old triangulation that intersect these grid squares are found.
The set of edges from these triangles that do not intersect these grid cells form a ring
around the area of our new triangles. The edges in this ring become constraints
in the triangulation. All of the points and constraint edges from the old triangulation are
included as well as the points and constraint edges from the new triangulation. Triangulating
these points and edges gives the new geometry for the modified area and has no effect on any
other part of the canvas. Therefore, the old triangles from the rest of the canvas can be reused,
and their union with the new triangle mesh represents the new canvas.

\section{Discussion and Results}
Paintings made with our system can be seen in Figures~\ref{fig:teaser}, \ref{fig:face},  \ref{fig:stop} and \ref{fig:boxes}.  Here we first consider how the complexity of the algorithm grows both in time and geometry.
The results from this analysis and our experience with the system have lead to a number of observations.

@@

Figure~\ref{fig:numtriangles} shows the zooming case grows linearly while the non-zooming case
appears to grow quadratically. This is expected since every new stroke has the chance to overlap
all of the old strokes. Therefore each new stroke does not just add a constant amount of geometry, but 
due to intersections with previous strokes extra intersection points are added. This results in a linear
increase in the amount of geometry per new stroke, yielding a quadratic growth rate overall. 
However, in the zooming case each stroke rarely overlaps previous strokes, and therefore
intersection points are rarely added. In this case the geometry increases by a constant amount per stroke, which yields a linear growth rate.

@@

\subsection{Stability}

While Shewchuk's implementation~\shortcite{shewchuk2002} of Delaunay triangulation is impressively robust, there are still certain pathological geometries that can cause it to fail.  Specifically, long skinny triangles and vertices that are too close together seem to trigger triangulation failures.  Detecting these situations and fixing them, or even better modifying the geometry construction to avoid them, would solve this problem.  For example, identifying if a new vertex is too close to an existing one, and if so, merging them.  Another example is to subdivide long edges to reduce the incidence of skinny triangles.  The impact these solutions would have on processing time is something that needs to be explored.

\begin{figure}
    \centering
        \fbox{\includegraphics[width=.45\textwidth]{images/grid}}
    \caption{An example of how our system determines which triangles to save and which to retriangulate. Blue
    lines represent the boundaries of old strokes, orange lines are the boundaries of the new stroke, red lines are
    the triangles that will be reused, and purple lines define the area that needs to be retriangulated. The purple
    lines are edges of triangles that intersect the new stroke. Since they interesect the new stroke they are no
    longer valid and must be retriangulated. All other triangles on the canvas can be reused.}
    \label{fig:modified}
\end{figure}

@@

\begin{figure}
  \centering
  \begin{subfigure}[b]{0.5\columnwidth}
    \centering
    % \includegraphics[width=\columnwidth]{graphs/numtriangles}
    \includegraphics[height=1.2in]{graphs/numtriangles}
    \caption{Number of triangles}
    \label{fig:numtriangles}
  \end{subfigure}%
  \begin{subfigure}[b]{0.5\columnwidth}
    \centering
    % \includegraphics[width=\columnwidth]{graphs/cumulativetime}
    \includegraphics[height=1.2in]{graphs/cumulativetime}
    \caption{Rendering time}
    \label{fig:timing}
  \end{subfigure}%
  \caption{For paintings of 50 random strokes, the left graph shows the number of triangles in the canvas as of stroke $x$, and the right graph shows the cumulative render time of the painting. The blue points are the non-zooming case, while the red points are the zooming case.}
\end{figure}

@@

% - Switch 6.3 and 6.4
\begin{figure*}
    \centering
        \fbox{\includegraphics[width=0.49\textwidth]{images/facezoom}}
        \fbox{\includegraphics[width=0.49\textwidth]{images/facezoomtriangles}}
    \caption{A drawing by an artist that took about 30 minutes, resulting in 9030 triangles.} % \Steve{Can we add number of strokes?}}
    \label{fig:face}
\end{figure*}

\begin{figure*}
    \centering
        \fbox{\includegraphics[width=0.325\textwidth]{images/stop12}}
        \fbox{\includegraphics[width=0.325\textwidth]{images/stop22}}
        \fbox{\includegraphics[width=0.325\textwidth]{images/stop32}}
    \caption{Another painting by an artist, using soft and hard strokes at a variety of scales.  Middle and right images are zoomed in.}
    \label{fig:stop}
\end{figure*}

\begin{figure*}
    \centering
        \includegraphics[width=0.195\textwidth]{images/zoom61}
        \includegraphics[width=0.195\textwidth]{images/zoom62}
        \includegraphics[width=0.195\textwidth]{images/zoom63}
        \includegraphics[width=0.195\textwidth]{images/zoom64}
        \includegraphics[width=0.195\textwidth]{images/zoom65}
        \includegraphics[width=0.195\textwidth]{images/zoom66}
        \includegraphics[width=0.195\textwidth]{images/zoom67}
        \includegraphics[width=0.195\textwidth]{images/zoom68}
        \includegraphics[width=0.195\textwidth]{images/zoom69}
        \includegraphics[width=0.195\textwidth]{images/zoom72}
    \caption{A painting that shows extreme zoom levels.  The final image is at zoom level 524,000:1, for an effective resolution equivalent to an image with more than $10^{17}$ pixels.}
    \label{fig:boxes}
\end{figure*}

\section{Conclusion}
Our system enables a user to paint like they would in a raster graphics program 
to create vector graphics. The core element is a triangle mesh representation of the painting.
We have created an algorithm to convert a user's mouse motions into a triangle mesh, and then composite that mesh on the preexisting mesh containing the previous strokes.

There are a number of intersting areas for future work, including:

%\noindent
%{\bf Simplification.}
%Our prototype implementation provides a baseline painting experience, in

Future work should focus on
implementing a simplification algorithm to allow the complexity of the drawings to grow. Other
improvements might focus on creating interesting manipulation tools, or other operations that utilize
the underlying triangle mesh structure to make the artists' job easier.

\ignorethis{
\section*{Acknowledgments}
Omitted for review.
}
